Pipeline Performance Figure Generation

### 2023-04-01 TestingPipeline/Scripts/paper_figure_generation.Rmd
Updating figures.

## Calculate the performance of each pipeline
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*10),
                      precision=rep(0, nrow(combos_list)*10),
                       recall=rep(0, nrow(combos_list)*10),
                       F1=rep(0, nrow(combos_list)*10),
                       MCC=rep(0, nrow(combos_list)*10), 
                      AUC=rep(0, nrow(combos_list)*10))

accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==1,], combos_list))
for (i in 2:10) {
  accuracy_scores <- rbind(accuracy_scores,
                           cbind(testing_set_index=rep(i, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==i,], combos_list)))
}
```

@@

@@

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Rule Sets",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rule Sets",
                     values = alpha(rev(pal(6)), 1))

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 1))
```
```{r}
mean_MCC <- accuracy_scores %>% 
  select(MCC, toolcombo) %>%
  group_by(toolcombo) %>%
  summarise(mean = mean(MCC))
```

```{r}
ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
                                  color=numrules, fill=numrules)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Score") +
  xlab("Number of Rule Sets") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1)) +
  facet_wrap(~performance_metric)

ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
                                  color=numtools, fill=numtools)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Score") +
  xlab("Number of Rules") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(pal(6)), 1)) +
  facet_wrap(~performance_metric)

ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numtools, y=performance_metric_score, 
                                  color=numrules, fill=numrules)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Score") +
  xlab("Number of Tools") +
  scale_fill_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="Number of Rules",
                     values = alpha(rev(pal(6)), 1)) +
  facet_wrap(~performance_metric)

@@

## visualizing a select set of tools
```{r}
viruses$keep_score_high_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)
viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_high"), size_class, true_virus) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, size_class, true_virus) %>%
  summarise(n = n())

```{r}
ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "PuOr", ) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~true_virus, scales = "free")
```

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_all_tv"), size_class, seqtype) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, size_class, seqtype) %>%
  summarise(n = n())

```{r}
pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_all_ntv"), size_class, seqtype) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, size_class, seqtype) %>%
  summarise(n = n())

```{r}
ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.7)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free")
```

@@

>>>>

## Calculate the performance of each pipeline
quantifying variability between rule sets
```{r}
accuracy_scores <- data.frame(testing_set_index=rep(0, nrow(combos_list)*8),
                      precision=rep(0, nrow(combos_list)*8),
                       recall=rep(0, nrow(combos_list)*8),
                       F1=rep(0, nrow(combos_list)*8),
                       MCC=rep(0, nrow(combos_list)*8), 
                      AUC=rep(0, nrow(combos_list)*8))

accuracy_scores <- cbind(testing_set_index=rep(1, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==1,], combos_list))
for (i in 2:8) {
  accuracy_scores <- rbind(accuracy_scores,
                           cbind(testing_set_index=rep(i, nrow(combos_list)),
                              build_score_list(viruses[viruses$Index==i,], combos_list)))
}
```

```{r}
MCC_sd_quant <- cbind(accuracy_scores$toolcombo[accuracy_scores$testing_set_index==1], as.data.frame(matrix(data=0, nrow=63, ncol=7)))

for (i in (2:8)) {
  temp <- accuracy_scores %>%
    filter(testing_set_index<=i) %>%
    select(MCC, toolcombo) %>%
    group_by(toolcombo) %>%
    summarise(sd = sd(MCC))

  MCC_sd_quant[,i] <- temp$sd
}

colnames(MCC_sd_quant) <- c("testing_set_index", "1-2", "1-3", "1-4",
                            "1-5", "1-6", "1-7", "1-8")

MCC_sd_quant <- MCC_sd_quant %>% pivot_longer(cols=contains("-"))

ggplot(MCC_sd_quant, aes(x=testing_set_index, y=value, 
                                  color=name, fill=name)) +
  geom_point(alpha=0.5)

ggplot(MCC_sd_quant, aes(x=name, y=value)) +
  geom_boxplot(alpha=0.5, color=pal(4)[1]) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 0),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Testing Sets") +
  ylab("Standard Deviation") 
```

## Calculate the performance of each pipeline
```{r}
accuracy_scores <- accuracy_scores %>% filter(testing_set_index<=5)
```

@@

### Visualize performance across all the rules

@@

fig <- ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numrules, fill=numrules)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 1))

ggsave(
  "../IntermediaryFiles/all_MCC_scatterplot.png",
  plot = fig,
  scale = 1,
  width = 8,
  height = 4,
  units = c("in", "cm", "mm", "px"),
  dpi = 300,
  limitsize = TRUE
)

ggplot(accuracy_scores, aes(x=toolcombo, y=MCC, 
                                  color=numtools, fill=numtools)) +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=20),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("MCC") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(viridis(6)), 1))
```

making heatmap for the x-axis label
```{r}
viruses_1 <- accuracy_scores %>% filter(testing_set_index==1)

ann_col = data.frame(
    tnv = combos_list$tune_not_viral,
    tv = combos_list$tune_viral,
    dvf = combos_list$DVF,
    vb = combos_list$VIBRANT,
    vs = combos_list$VS,
    vs2 = combos_list$VS2
)

ann_colors = list(
  tnv = c("black", pal(6)[1]),
  tv = c("black", pal(6)[3]),
  dvf = c("black", pal(6)[2]),
  vb = c("black", pal(6)[4]),
  vs = c("black", pal(6)[5]),
  vs2 = c("black", pal(6)[6])
)

rownames(ann_col) <- combos_list$toolcombo

viral_scores_1 <- accuracy_scores %>% filter(testing_set_index==1)

rownames(viral_scores_1) <- viral_scores_1$toolcombo

viral_scores_1 <- viral_scores_1 %>%
  select(c("MCC"))

pheatmap::pheatmap(as.matrix(viral_scores_1),
                  annotation_row = ann_col,
                  annotation_colors = ann_colors,
                  border_color = NA,
                  show_rownames = F,
                  show_colnames = F,
                  annotation_legend = F,
                  cluster_rows = F,
                  cluster_cols = F,
                  cellheight=3,
                  filename = "../IntermediaryFiles/heatmap_MCC_labels.png"
                  )
```

```{r}
mean_MCC <- accuracy_scores %>% 
  select(MCC, toolcombo) %>%
  group_by(toolcombo) %>%
  summarise(mean = mean(MCC))
```

```{r}
ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
                                  color=numrules, fill=numrules)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Score") +
  xlab("Number of Rule Sets") +
  scale_fill_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 1)) +
  facet_wrap(~performance_metric)

ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numrules, y=performance_metric_score, 
                                  color=numtools, fill=numtools)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Score") +
  xlab("Number of Rules") +
  scale_fill_manual(name="Number of Tools",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Tools",
                     values = alpha(rev(viridis(6)), 1)) +
  facet_wrap(~performance_metric)

ggplot(accuracy_scores_melt[accuracy_scores_melt$performance_metric!="MCC",], aes(x=numtools, y=performance_metric_score, 
                                  color=numrules, fill=numrules)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Score") +
  xlab("Number of Tools") +
  scale_fill_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 0.5)) +
  scale_color_manual(name="Number of Rule Sets",
                     values = alpha(rev(viridis(6)), 1)) +
  facet_wrap(~performance_metric)

@@

## visualizing a select set of tools
```{r}
viruses$keep_score_high_all <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
viruses$keep_score_high_recall <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)
viruses$keep_score_high_precision <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
viruses$keep_score_high_MCC <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_high"), true_virus) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, true_virus) %>%
  summarise(n = n())

pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~true_virus, scales = "free")
```

```{r}
viruses_long_scores <- viruses %>% 
  select(contains("keep_score_high"), max_score_group, true_virus) %>%
  filter(true_virus=="virus") %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, max_score_group) %>%
  summarise(n = n())

pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
pal_orange <- RColorBrewer::brewer.pal(3, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~max_score_group, scales = "free") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_high"), max_score_group, true_virus) %>%
  filter(true_virus=="not") %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, max_score_group) %>%
  summarise(n = n())

pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom"
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~max_score_group, scales = "free") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
```

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_all_tv"), seqtype) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, seqtype) %>%
  summarise(n = n())

pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom",
    axis.text.x=element_text(size=8, angle = 45),
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
```

@@

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_all_ntv"), seqtype) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, seqtype) %>%
  summarise(n = n())
```

```{r}
ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom",
    axis.text.x=element_text(size=8, angle = 45),
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 2))
```

@@

comparing each tool's score
```{r}
viruses$keep_score_ind_vs2 <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_ind_tv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_ind_dvf <- getting_viral_set_1(viruses, include_deepvirfinder = T,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_ind_vb <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = T,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = F)

viruses$keep_score_ind_ntv <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)

viruses$keep_score_ind_vs <- getting_viral_set_1(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = F,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)

viruses_long_scores <- viruses %>% 
  select(contains("keep_score_ind_"), seqtype) %>%
  pivot_longer(cols=contains("keep_score_"), 
               names_to="rule_combination",
               values_to="viral_score") %>% 
  mutate(viral_score=as.factor(round(viral_score))) %>%
  group_by(rule_combination, viral_score, seqtype) %>%
  summarise(n = n())
```
Fig 4

```{r}
num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)

for (i in 1:63) {
  for (j in 1:63) {
    num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)

    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
  }
}

rownames(num_high_viruses) <- combos_list$toolcombo
colnames(num_high_viruses) <- combos_list$toolcombo

rownames(prop_high_viruses) <- combos_list$toolcombo
colnames(prop_high_viruses) <- combos_list$toolcombo
```

```{r}
pal_purple <- RColorBrewer::brewer.pal(6, "Purples")
pal_orange <- RColorBrewer::brewer.pal(4, "Oranges")

ggplot(viruses_long_scores, aes(y=n, x=rule_combination,
                   fill=viral_score)) +
  geom_bar(stat="identity", color="black") +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom",
    axis.text.x=element_text(size=8, angle = 45)
  ) +
  scale_fill_manual(name="",
                     values = alpha(c(rev(pal_orange), pal_purple), 0.8)) +
  xlab("") +
  ylab("Number of Sequences") + 
  facet_grid(~seqtype, scales = "free") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 2)) +
  scale_x_discrete(labels=rev(c("VirSorter2", "VirSorter", "VIBRANT",
                            "tuning viral", "tuning not viral",
                            "DeepVirFinder")))
```

# Comparing all of the tool methods - Fig 2A
```{r}
combos_list <- data.frame(toolcombo=rep(0, 64),
                          tune_not_viral=rep(0, 64),
                          DVF=rep(0, 64),
                          tune_viral=rep(0, 64),
                          VIBRANT=rep(0, 64),
                          VS=rep(0, 64),
                          VS2=rep(0, 64))
p <- 1

for (i in c(0,1)){
  for (j in c(0,1)){
    for (k in c(0,1)){
      for (l in c(0,1)){
        for (m in c(0,1)){
          for (n in c(0,1)){
            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
            combos_list$toolcombo2[p] <- paste(if(i){"tnv"}else{"0"},if(j){"DVF"}else{"0"},
                                               if(k){"tv"}else{"0"},if(l){"VB"}else{"0"},
                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
            combos_list$tune_not_viral[p] <- i
            combos_list$DVF[p] <- j
            combos_list$tune_viral[p] <- k
            combos_list$VIBRANT[p] <- l
            combos_list$VS[p] <- m
            combos_list$VS2[p] <- n
            p <- p+1
          }
        }
      }
    }
  }
}

combos_list <- combos_list[-1,]

viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
                          num_viruses=rep(0, nrow(combos_list)))

for (i in 1:nrow(combos_list)) {
  viral_scores[,i] <- getting_viral_set_1(viruses, include_vibrant = combos_list$VIBRANT[i],
                                            include_virsorter = combos_list$VS[i],
                                            include_virsorter2 = combos_list$VS2[i],
                                            include_tuning_viral = combos_list$tune_viral[i],
                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
                                            include_deepvirfinder = combos_list$DVF[i])

  if (max(viral_scores[,i])<=0) {
    num_viruses$num_viruses[i] <- 0
  }
  else {
    num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
  }
  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]

  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
}
num_viruses$numtools <- str_count(num_viruses$toolcombo, "1")
num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
num_viruses$numtools <- as.factor(num_viruses$numtools)

pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")

viruses_1 <- viruses[viruses$Index=="1",]

ann_col = data.frame(
    tnv = combos_list$tune_not_viral,
    tv = combos_list$tune_viral,
    dvf = combos_list$DVF,
    vb = combos_list$VIBRANT,
    vs = combos_list$VS,
    vs2 = combos_list$VS2
)

ann_row = data.frame(
    score = viruses_1$seqtype
)

ann_colors = list(
  tnv = c("black", pal(6)[1]),
  tv = c("black", pal(6)[3]),
  dvf = c("black", pal(6)[2]),
  vb = c("black", pal(6)[4]),
  vs = c("black", pal(6)[5]),
  vs2 = c("black", pal(6)[6]),
  score = c("archaea"=pal(6)[1], "bacteria"=pal(6)[2], "fungi"=pal(6)[3],
              "plasmid"=pal(6)[4], "protist"=pal(6)[5], "virus"=pal(6)[6])
)

rownames(ann_col) <- combos_list$toolcombo
rownames(ann_row) <- viruses_1$contig

viral_scores_1 <- viral_scores[viruses$Index=="1",]

colnames(viral_scores_1) <- combos_list$toolcombo
rownames(viral_scores_1) <- viruses_1$contig

pheatmap::pheatmap(t(viral_scores_1),
                  annotation_col = ann_row,
                  annotation_row = ann_col,
                  annotation_colors = ann_colors,
                  #border_color = "black",
                  color = c(rev(RColorBrewer::brewer.pal(3,"Oranges")),
                            RColorBrewer::brewer.pal(5,"Purples") 
                            ),
                  show_rownames = F,
                  show_colnames = F,
                  annotation_legend = F,
                  cluster_rows = F,
                  cluster_cols = F,
                  cellheight=3,
                  filename = "../IntermediaryFiles/pheatmap_virus_set_1_viral_scores.png"
                  )
```

# comparing all rules against each other - Figure 5
```{r}
num_high_viruses <- matrix(data=NA, nrow=63, ncol=63)
prop_high_viruses <- matrix(data=NA, nrow=63, ncol=63)

for (i in 1:63) {
  for (j in 1:63) {
    num_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)

    prop_high_viruses[i,j] <- sum(viral_scores[,i]>=1 & viral_scores[,j]>=1)
    prop_high_viruses[i,j] <- prop_high_viruses[i,j]/sum(viral_scores[,i]>=1 | viral_scores[,j]>=1)
  }
}

rownames(num_high_viruses) <- combos_list$toolcombo
colnames(num_high_viruses) <- combos_list$toolcombo

rownames(prop_high_viruses) <- combos_list$toolcombo
colnames(prop_high_viruses) <- combos_list$toolcombo
```

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")

ann_col = data.frame(
    tnv = combos_list$tune_not_viral,
    tv = combos_list$tune_viral,
    dvf = combos_list$DVF,
    vb = combos_list$VIBRANT,
    vs = combos_list$VS,
    vs2 = combos_list$VS2,
    num_vir = colSums(viral_scores>=1)+1
)

ann_colors = list(
  num_vir=magma(100),
  tnv = c("black", pal(6)[1]),
  tv = c("black", pal(6)[3]),
  dvf = c("black", pal(6)[2]),
  vb = c("black", pal(6)[4]),
  vs = c("black", pal(6)[5]),
  vs2 = c("black", pal(6)[6])
)

rownames(ann_col) <- combos_list$toolcombo

pheatmap::pheatmap(prop_high_viruses,
                  annotation_row = ann_col,
                  annotation_colors = ann_colors,
                  border_color = "black",
                  color = viridis(100),
                  cutree_rows = 3,
                  cutree_cols = 3,
                  show_rownames = F,
                  show_colnames = F,
                  filename = "../IntermediaryFiles/pheatmap_prop_viruses_in_common.png",
                  annotation_legend = F)
```
